
# A TIME SERIES INVESTIGATION OF HARMFUL AIR POLLUTANTS CONCENTRATIONS IN BERLIN
## By Subhamathi Thiruvenkatam
### 2023-03-09


# Introduction

## General Background:
>Air pollution is a major environmental issue in Germany, and the country has taken several measures to monitor and reduce its air pollution levels. Ozone (O3), nitrogen dioxide (NO2), and nitrogen oxide (NO) are some of the important air pollutants that contribute majorly towards the pollution sector. NO2 is a toxic gas that is largely released by diesel automobiles and is likely to cause health problems related to heart and lung disorders. Vehicles and power plants both emit NO, which is a precursor to NO2. O3 is a secondary pollutant that can destroy plant and animal life as well as cause respiratory issues.Monitoring for air pollution levels helps in locating the sources of the pollution, such as manufacturing facilities, power plants, and vehicles. These sources can be controlled and their emissions can be decreased using this knowledge.Germany has implemented various initiatives in the recent years to control the air pollution and improve the air quality

## Aim and Objective:
>The main aim of this research paper is to monitor Berlin's air pollution levels and analyse the changes in their patterns with the help of a time series plot. The air pollution levels of the three main air pollutants , NO2,NO,03 are monitored for this study.The hypotheses is taken as the air pollutant NO2 is below the threshold set and the three pollutants are correlated to each other.

# Study Area:
> Germany's capital and largest city, Berlin, is situated in the northeastern region of the nation. The city, which has a population of over 3.7 million, is renowned for its rich cultural past, and diverse neighbourhoods.Berlin is a significant metropolitan center that serves as a focus for commerce and transportation, resulting in higher levels of air pollution. Berlin is an important area to study the air pollutants characteristics as it is rich in socio-economic activities, population and industrial sectors. 

# Datasets Used:
>1. The annual mean NO2 data for Berlin
>2. The annual mean O3 data for Berlin
>3. The annual mean NO data for Berlin

## Data Source:
>https://www.eea.europa.eu/data-and-maps/data/aqereporting-9

# Methodology:
## Libraries Used:
>1.tidyverse for tiding up the data

>2.lubridate for formatting the data

>3.readr for reading csv files

>4.ggplot2 for data visualization

>5.reshape2 for melting the correlation matrix

>6.dplyr for data manipulation

>7.corrplot for correlation plots

## Method:
>The research starts with preparing the data and importing the data using the read.csv() function.Since I am planned to analyse the data from a single station , the data imported is then filtered to contain only air pollution levels from the station Berlin Neukolln which has 10 years data for all three pollutants. The data is imported in the same way as for NO2,03,NO. 
> The tidyverse library along with ggplot() is used to plot the location of the station on the Berlin city 
> The 10 years data for NO2 is binded into a single file df1, for 03 and NO the same process is repeated using the bind_rows() function
>All the air pollutants data is binded into one for the final analysis
>Using ggplot() the air pollutants levels across the Years are plotted as a line graph and all the plots are combined into a single grid 
>The correlation matrix is created for these pollutants and displayed using a heatmap indicating their correlation coefficients using ggplot() and melt()

# Data preparation:
```{r echo=TRUE}
library(tidyverse)
library(lubridate)
library(readr)
library(ggplot2)
library(dplyr)
library(corrplot)
library(ggplot2)
library(reshape2)
#Data for NO2
data2010 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2010/b2010.csv")
d10 <- data2010[data2010$Air.Quality.Station.Name == "Berlin Neukolln",]
data2011 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2011/b2011.csv")
d11 <- data2011[data2011$Air.Quality.Station.Name == "Berlin Neukolln",]
data2012 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2012/b2012.csv")
d12 <- data2012[data2012$Air.Quality.Station.Name == "Berlin Neukolln",]
data2013 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2013/b2013.csv")
d13 <- data2013[data2013$Air.Quality.Station.Name == "Berlin Neukolln",]
data2014 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2014/b2014.csv")
d14 <- data2014[data2014$Air.Quality.Station.Name == "Berlin Neukolln",]
data2015 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2015/b2015.csv")
d15 <- data2015[data2015$Air.Quality.Station.Name == "Berlin Neukolln",]
data2016 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2016/b2016.csv")
d16 <- data2016[data2016$Air.Quality.Station.Name == "Berlin Neukolln",]
data2017 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2017/b2017.csv")
d17 <- data2017[data2017$Air.Quality.Station.Name == "Berlin Neukolln",]
data2018 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2018/b2018.csv")
d18 <- data2018[data2018$Air.Quality.Station.Name == "Berlin Neukolln",]
data2019 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2019/b2019.csv")
d19 <- data2019[data2019$Air.Quality.Station.Name == "Berlin Neukolln",]
data2020 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2020/b2020.csv")
d20 <- data2020[data2020$Air.Quality.Station.Name == "Berlin Neukolln",]



#Data for O3
O3_2010 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2010/O3_10.csv")
o10 <- O3_2010[O3_2010$Air.Quality.Station.Name == "Berlin Neukolln",]
O3_2011 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2011/O3_11.csv")
o11 <- O3_2011[O3_2011$Air.Quality.Station.Name == "Berlin Neukolln",]
O3_2012 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2012/O3_12.csv")
o12 <- O3_2012[O3_2012$Air.Quality.Station.Name == "Berlin Neukolln",]
O3_2013 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2013/O3_13.csv")
o13 <- O3_2013[O3_2012$Air.Quality.Station.Name == "Berlin Neukolln",]
O3_2014 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2014/O3_14.csv")
o14 <- O3_2014[O3_2014$Air.Quality.Station.Name == "Berlin Neukolln",]
O3_2015 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2015/O3_15.csv")
o15 <- O3_2015[O3_2015$Air.Quality.Station.Name == "Berlin Neukolln",]
O3_2016 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2016/O3_16.csv")
o16 <- O3_2016[O3_2016$Air.Quality.Station.Name == "Berlin Neukolln",]
O3_2017 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2017/O3_17.csv")
o17 <- O3_2017[O3_2017$Air.Quality.Station.Name == "Berlin Neukolln",]
O3_2018 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2018/O3_18.csv")
o18 <- O3_2018[O3_2018$Air.Quality.Station.Name == "Berlin Neukolln",]
O3_2019 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2019/O3_19.csv")
o19 <- O3_2019[O3_2019$Air.Quality.Station.Name == "Berlin Neukolln",]
O3_2020 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2020/O3_20.csv")
o20 <- O3_2020[O3_2020$Air.Quality.Station.Name == "Berlin Neukolln",]




#Data for NO
NO_2010 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2010/NO_10.csv")
NO10 <- NO_2010[NO_2010$Air.Quality.Station.Name == "Berlin Neukolln",]
NO_2011 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2011/NO_11.csv")
NO11 <- NO_2011[NO_2011$Air.Quality.Station.Name == "Berlin Neukolln",]
NO_2012 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2012/NO_12.csv")
NO12 <- NO_2012[NO_2012$Air.Quality.Station.Name == "Berlin Neukolln",]
NO_2013 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2013/NO_13.csv")
NO13 <- NO_2013[NO_2012$Air.Quality.Station.Name == "Berlin Neukolln",]
NO_2014 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2014/NO_14.csv")
NO14 <- NO_2014[NO_2014$Air.Quality.Station.Name == "Berlin Neukolln",]
NO_2015 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2015/NO_15.csv")
NO15 <- NO_2015[NO_2015$Air.Quality.Station.Name == "Berlin Neukolln",]
NO_2016 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2016/NO_16.csv")
NO16 <- NO_2016[NO_2016$Air.Quality.Station.Name == "Berlin Neukolln",]
NO_2017 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2017/NO_17.csv")
NO17 <- NO_2017[NO_2017$Air.Quality.Station.Name == "Berlin Neukolln",]
NO_2018 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2018/NO_18.csv")
NO18 <- NO_2018[NO_2018$Air.Quality.Station.Name == "Berlin Neukolln",]
NO_2019 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2019/NO_19.csv")
NO19 <- NO_2019[NO_2019$Air.Quality.Station.Name == "Berlin Neukolln",]
NO_2020 <- read.csv("C:/Users/subhu/Desktop/AOSD/New folder/berlin no2/2020/NO_20.csv")
NO20 <- NO_2020[NO_2020$Air.Quality.Station.Name == "Berlin Neukolln",]

``` 

### Binding all the NO2 data 
```{r echo=TRUE}
df1 <- bind_rows(d10, d11, d12, d13, d14, d15, d16, d17, d18, d19, d20)
Nitrogen_dioxide <- df1$Air.Pollution.Level
df1
```

### Binding all the O3 Data
```{r echo=TRUE}
O3 <- bind_rows(o10, o11, o12, o13, o14, o15, o16, o17, o18, o19, o20)
Ozone <- O3$Air.Pollution.Level
O3
```

### Binding all the NO data
```{r echo=TRUE}
NO <- bind_rows(NO10, NO11, NO12, NO13, NO14, NO15, NO16, NO17, NO18, NO19, NO20)
Nitrogen_Monoxide <- NO$Air.Pollution.Level
NO
```

### Combining all the Air pollutants data
```{r echo=TRUE}
All <- bind_rows(df1,O3,NO)
All
```

### Plotting the station chosen 

```{r echo=TRUE}
library(tidyverse) |> suppressPackageStartupMessages()
library(sf)

crs <- st_crs("EPSG:32632")
st_as_sf(d12, crs = "OGC:CRS84", coords = 
           c("Longitude", "Latitude")) |>
  st_transform(crs) -> no2.sf

read_sf("C:/Users/subhu/Desktop/AOSD/New folder/Berlin/Berlin_city.shp") |> st_transform(crs) -> de


ggplot() + geom_sf(data = de) + 
  geom_sf(data = no2.sf)

```

### Analysis

### NO2 data 

```{r echo=TRUE}
# Visualizing the plot for NO2
options(repr.plot.width = 16, repr.plot.height = 6)

ggplot(data = df1, aes(x = Year, y = Air.Pollution.Level, color = Air.Pollution.Level)) +
  geom_line() +
  scale_color_gradient(low = "green", high = "red") +
  labs(x = "Year", y = "Concentration (ug/m3)", title="Trends in NO2 Concentrations Over Time in Berlin ")

```


> From the above graph we can see that the line graph shows a consistent decrease in NO2 levels over a 10-year period. The x-axis represents time and the y-axis represents NO2 levels in ug/m3. The color gradient ranges from red to green indicating the red lines to be higher values and green lines to be the lower ones.

>The trend line shows a gradual decline in NO2 levels from 2010 to 2016 but a steep decline from 2016 till 2020. This may account to the taken by the Government to reduce air pollution by electrifying transportation and retrofitting diesel buses.These actions maybe the influence of the " Immediate Action Programme for Clean Air " that ran from 2017 to 2020.


### Ozone data

```{r echo=TRUE}
# Visualizing the plot for O3
ggplot(data = O3, aes(x = Year, y = Air.Pollution.Level)) +
  geom_line(color="Blue") +
  labs(x = "Year", y = "Concentration (ug/m3)",title="Trends in Ozone Concentrations Over Time in Berlin ")

```

> It can be observed from the graph that the Ozone values do not have a stable pattern. There are crests and troughs throughout the 10 year period.The x-axis represents time and the y-axis represents NO2 levels in ug/m3

>Unlike Particulate matter and Nitrogen dioxide , ozone is not emitted directly but formed from NO and organic compounds. Solar radiation is also a main reason for this. Thus when there are long term high pressure summer days, ozone can be stored in the lower layers of the atmosphere leading to its heavy concentration. Therefore the cause for the variation of the levels throughtout the period maybe due to the weather conditions.



### NO data

```{r echo=TRUE}
# Visualizing the plot for NO
ggplot(data = NO, aes(x = Year, y = Air.Pollution.Level, color = Air.Pollution.Level)) +
  geom_line() +
  scale_color_gradient(low = "PINK", high = "VIOLET") +
  labs(x = "Year", y = "Concentration (ug/m3)", title="Trends in NO Concentrations Over Time in Berlin ")
```

>The x-axis represents time and the y-axis represents NO2 levels in ug/m3. The light pink color indicates the lowest value and increase in the color shows the higher levels.

>The NO graph shows a decline in the concentrations over the period of 10 years with a steep decline in 2013 and a raise after that. There is a steady decrease from 2015 to 2020. 


# Combining all the plots into one grid
```{r echo=TRUE}
# Combine the plots into one grid

df <- data.frame(x = All$Year,
                 y = All$Air.Pollution.Level,
                 Air.Pollutants = factor(All$Air.Pollutant.Description))


ggplot(df, aes(x = x, y = y, color = Air.Pollutants)) + 
  geom_line() + 
  labs(x = "Time (years) ", y = "Concentration (ug/m3)", title = "Decadal Changes in Harmful Air Pollutants Concentrations" )+
  scale_color_manual(values = c("red", "green", "blue"))


```

> It can be seen that the Ozone levels are harming compared to the NO2 and NO which lie below to it. The NO and NO2 levels have a stable decline for the 10 year time period chosen whereas the Ozone has high variations in the peak concentrations.


# Correlation Matrix

> A dataframe is created with the values of NO2,O3,NO . The correlation matrix is created for the three pollutants using the melt() and visualized using the heatmap

```{r echo=TRUE}
# Creating a data frame with the three pollutants 
pollutants_df <- data.frame(Nitrogen_dioxide, Ozone, Nitrogen_Monoxide)

# calculation of the correlation matrix
cor_matrix <- cor(pollutants_df)
cor_matrix



#Visualizing the correlation matrix
cor_df <- melt(cor_matrix)


# creation of a heatmap
ggplot(cor_df, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "white", mid = "grey", high = "black", 
                       midpoint = 0, limit = c(-3,3)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0))

```

# Results :

> The air quality standard set by the European Union for  NO2 on annual mean basis is 40 ug/m3. From the time series plot measured it can be seen that the NO2 values are under the threshold and comply within the EU standards making the air quality much better and sustainable.

> The peak concentration of O3 at 2019 is an alarming one and the NO levels are similar to that of the NO2 and are not as high as Ozone levels

> The correlation matrix results that the Pollutant NO2 is positively high correlated with NO and is negatively correlated with O3 and Ozone is negatively correlated with NO 


# Conclusion and Discussion :

> The overall analyses suggests that the policies and initiatives taken by the German Government is effective and the pollution levels have decreased improving the air quality to a greater extent. The transportational changes and bio changes have had a great impact on the environment.

>The NO2 levels are well under the standard in this station suggesting that this part of the city is well formed and adapted.

> The Ozone levels influenced by the weather conditions suggest that they are to be monitored at a greater level for changes. Sudden increase in their levels are not because of the direct emission indicating that they need to be monitored at a higher temporal scale.


# References:

>https://environment.ec.europa.eu/topics/air/air-quality/eu-air-quality-standards_en

>Kurtenbach, R., Kleffmann, J., Niedojadlo, A. et al. Primary NO2 emissions and their impact on air quality in traffic environments in Germany. Environ Sci Eur 24, 21 (2012)

>Hagenbjörk, A., Malmqvist, E., Mattisson, K. et al. The spatial variation of O3, NO, NO2 and NO x and the relation between them in two Swedish cities. Environ Monit Assess 189, 161 (2017).

>https://r-spatial.org/book/12-Interpolation.html






